Temporal ordering and registration of images in studies of developmental dynamics

### 2015-02-02 image_analysis_paper_development2/SI_text.tex
updated report

For our general purposes, the relevant image operations are 
\begin{enumerate}
\item {\bf Intensity normalization:} Contrast-limited adaptive histogram equalization (using the \texttt{adapthisteq} function with an $8 \times 8$ tile grid, and a uniform distribution for the intensities with a clip limit of 0.01) is used to normalize the intensities of signals whose absolute intensity is not meaningful/informative.
%
\item {\bf Blur:} The \texttt{imfilter} function with a disc filter is used to blur signals whose small-scale structure is not informative. For example, 
%
\item {\bf Intensity scaling (multichannel images only):} The \texttt{immultiply} function is used to scale signal intensities. This is important for multichannel images, and it determines the (relative) contributions of the various signals.  
%
\item {\bf Mean-centering:}  The Canny method \citep{canny1986computational} is used to detect the edges of the object in each image (using the \texttt{edge} function). The image is then translated so that the object (as determined by extremeties of the detected edges) is centered.
%
\item {\bf Size scaling:} For images whose relative size is unimportant to the developmental dynamics, the images are rescaled/dialated so that the object size (as determined by extremeties of the detected edges) is constant (we set this to be 80\% of the total image).
%
\item {\bf Removing corners:} The image is cropped/filtered using a disc centered in the middle of the image and whose diameter is equal to the number of pixels; any pixels outside of this disc are set to 0 (no intensity). This removes any corner effects under rotations. 
\end{enumerate}

@@

\begin{table}[t]
%\caption{Outline of algorithm used to temporally order images}
\setlength{\fboxrule}{2pt}
\framebox[\columnwidth]{
\begin{minipage}{0.9\columnwidth}
{\bf Outline of algorithm used to register and order images}
\begin{enumerate}
\item Compute the alignments for each pair of images, as described in \eqref{eq:opt_pairwise}.
Retain both the optimal pairwise rotations as well as the minimum distances obtained when pairs are optimally aligned.
%
\item Compute the corresponding rotation matrices $R(\theta_{ij})$ from \eqref{eq:R_theta}.
%
\item Select $\epsilon$ for use in the diffusion maps kernel. 
%
Taking $\epsilon$ to be on the order of the median of the pairwise distances often yields good results. 
\item Compute the matrix $S$ in \eqref{eq:vdm_S}. 
$A_{ij}$ is computed from \eqref{eq:dmaps_W} and \eqref{eq:dmaps_A}, using the pairwise distances from step 1, and $H_{ij} = R(\theta_{ij})$ are computed in step 2.
%
\item Compute the eigenvalues  $\lambda_1, \lambda_2, \dots, \lambda_{2m}$ and eigenvectors $\phi_1, \phi_2, \dots, \phi_{2m}$ of $S$, ordered such that $|\lambda_1| \ge |\lambda_2| \ge \dots \ge |\lambda_{2m}|$.
%
\item Estimate the optimal three-dimensional rotation for each image: stack the first two eigenvectors $\phi_1, \phi_2$ into a $2m \times 2$ matrix, and then divide this matrix into $m$ $2 \times 2$ blocks, $\hat{R}_1, \dots, \hat{R}_m$. Compute the estimated rotations for each of these $m$ blocks as in \eqref{eq:R_est}. If most of the rotations are improper ($det(R) = -1$), flip the sign of one of the eigenvectors.
\item Multiply each estimated rotation $R_{i, est}$ by $R_{1, est}^T$ (so that $R_{1, est}$ will become the identity matrix).
\item For each image $i$, compute the optimal rotation to align the image by converting $R_{i, est}^T$ (note the transpose) to the corresponding angle using \eqref{eq:R_theta}. 
\item Compute the first embedding coordinate $\psi_{k,1}(i)$ as described in \eqref{eq:vdm_coord} and \eqref{eq:first_embed_vdm}, where $3 \le k \le 4$.
\item To order the images, sort them by the values of this embedding coordinate $\psi_{k,l}(i)$.
\end{enumerate} 
\end{minipage}
}
\end{table}

>>>>

For our general purposes, the relevant image operations are 
\begin{description}

\item[Intensity normalization] Contrast-limited adaptive histogram equalization (using the \texttt{adapthisteq} function with an $8 \times 8$ tile grid, and a uniform distribution for the intensities with a clip limit of 0.01) is used to normalize the intensities of signals whose absolute intensity is not meaningful/informative.
%
\item[Blur] The \texttt{imfilter} function with a disc filter is used to blur signals whose small-scale structure is not informative. For example, 
%
\item[Intensity scaling (multichannel images only)] The \texttt{immultiply} function is used to scale signal intensities. This is important for multichannel images, and it determines the (relative) contributions of the various signals.  
%
\item[Mean-centering]  The Canny method \citep{canny1986computational} is used to detect the edges of the object in each image (using the \texttt{edge} function). The image is then translated so that the object (as determined by extremeties of the detected edges) is centered.
%
\item[Size scaling] For images whose relative size is unimportant to the developmental dynamics, the images are rescaled/dialated so that the object size (as determined by extremeties of the detected edges) is constant (we set this to be 80\% of the total image).
%
\item[Removing corners] The image is cropped/filtered using a disc centered in the middle of the image and whose diameter is equal to the number of pixels; any pixels outside of this disc are set to 0 (no intensity). This removes any corner effects under rotations. 
\end{description}

@@

\begin{table}[t]
%\caption{Outline of algorithm used to temporally order images}
\setlength{\fboxrule}{2pt}
\framebox[\columnwidth]{
\begin{minipage}{0.9\columnwidth}
{\bf Outline of algorithm used to register and order images}
\begin{enumerate}
\item Compute the alignments for each pair of images, as described in \eqref{eq:opt_pairwise}.
Retain both the optimal pairwise rotations as well as the minimum distances obtained when pairs are optimally aligned.
%
\item Compute the corresponding rotation matrices $R(\theta_{ij})$ from \eqref{eq:R_theta}.
%
\item Select $\epsilon$ for use in the diffusion maps kernel. 
%
Taking $\epsilon$ to be $25-50\%$ of the median of the pairwise distances often yields good results. 
\item Compute the matrix $S$ in \eqref{eq:vdm_S}. 
$A_{ij}$ is computed from \eqref{eq:dmaps_W} and \eqref{eq:dmaps_A}, using the pairwise distances from step 1, and $H_{ij} = R(\theta_{ij})$ are computed in step 2.
%
\item Compute the eigenvalues  $\lambda_1, \lambda_2, \dots, \lambda_{2m}$ and eigenvectors $\phi_1, \phi_2, \dots, \phi_{2m}$ of $S$, ordered such that $|\lambda_1| \ge |\lambda_2| \ge \dots \ge |\lambda_{2m}|$.
%
\item Estimate the optimal three-dimensional rotation for each image: stack the first two eigenvectors $\phi_1, \phi_2$ into a $2m \times 2$ matrix, and then divide this matrix into $m$ $2 \times 2$ blocks, $\hat{R}_1, \dots, \hat{R}_m$. Compute the estimated rotations for each of these $m$ blocks as in \eqref{eq:R_est}. If most of the rotations are improper ($det(R) = -1$), flip the sign of one of the eigenvectors.
\item Multiply each estimated rotation $R_{i, est}$ by $R_{1, est}^T$ (so that $R_{1, est}$ will become the identity matrix).
\item For each image $i$, compute the optimal rotation to align the image by converting $R_{i, est}^T$ (note the transpose) to the corresponding angle using \eqref{eq:R_theta}. 
\item Compute the first embedding coordinate $\psi_{k,1}(i)$ as described in \eqref{eq:vdm_coord} and \eqref{eq:first_embed_vdm}, where $3 \le k \le 4$.
\item To order the images, sort them by the values of this embedding coordinate $\psi_{k,l}(i)$.
\end{enumerate} 
\end{minipage}
}
\end{table}

### 2015-02-02 image_analysis_paper_development2/code_for_distribution/README/README.tex
updated report

\usepackage{epstopdf}

@@

All algorithms and analysis were implemented in MATLAB\textsuperscript{\textregistered} (R2013b, The MathWorks, Natick, Massachusetts).
%
The code requires MATLAB along with the Image Processing Toolbox.

@@

There are two ways to use the code
\begin{enumerate}
\item Interactive GUI.
\item Program a simple script.
\end{enumerate}

The basic steps are as follows:
\begin{enumerate}
   \item Read in images
   \item Preprocess the images (smooth and/or equalize channels, mean-center, etc.)
   \item Run vector diffusion maps to register and order preprocessed images
   \item (Optional) Save registered and ordered images
   \item (Optional) Compute an average trajectory
\end{enumerate}

@@

@@

There are five steps
%
\begin{enumerate}
\item Read Images
\item Preprocess Images
\item Analyze Images
\item Save Images
\item Average Trajectory
\end{enumerate}

@@

The fields are as follows
\begin{description}
%
\item[2D/3D] Select ``2D'' if the images are traditional 2D images. Select ``3D'' if the images are z-stacks.
%
\item[Image Directory] Enter the directory where the images or z-stacks are stored. This directory can be an absolute path (e.g. \texttt{C:/Documents/Images}) or a path relative to the directory where this code is stored (e.g. \texttt{../Images})
%
\item[Image Prefix] The prefix for all image file names. The images are assumed to named as the image prefix, followed by a number. 
%
\item[Stack Prefix] (only for 3D z-stacks) The prefix for all stack folder names. Each stack is assumed to be stored in a single folder, with the folder's name as the stack prefix, followed by a number. 
%
\item[Image Extension] The extension of the images (e.g., tif, jpg)
%
\item[Number of Images] The number of images or z-stacks. Images are assumed to be stored as the image prefix, followed by a two-digit index, where the indices run from 1 to the number of images (e.g., image\_prefix01.tif, image\_prefix02.tif, etc.). For three-dimensional images, the z-stacks are assumed to be stored in individual folders whose names are the stack prefix, followed by a two-digit index, where the indices run from 1 to the number of stacks (e.g., z\_stack\_prefix01/, z\_stack\_prefix02/, etc.)
%
\item[Number of Images in Stack] (only for z-stacks) The number of images per z-stack. Images are assumed to be stored in the individual stack directories as the stack prefix, followed by a two-digit index, where the indices run from 1 to the number of images in the stack (e.g., stack\_prefix01.tif, stack\_prefix02.tif, etc.). 
%
\end{description}

@@

\begin{description}
\item[Weight] Each channel in the images is scaled by the weight factor, where all weights at 1 does not adjust the channel intensities. For grayscale images, only the first weight is used. 
%
\item[Blur] Each channel in the images is blurred using a disc factor with radius equal to a fraction of the image as indicated by the ``blur'' field. For example, setting the blur to 5\% means that the image is blurred using a disc filter with radius $0.05 \times$ the number of pixels. 
%
\item[Normalize] If the normalize option is selected for a specific channel, then Contrast-Limited Adaptive Histogram Equalization is applied to that channel to normalize the intensities across the image. This is important for signals whose absolute intensity is not meaningful or informative. 
%
\item[Use to Mean-Center] If the mean-center option is selected for any of the channels, then the object is centered in the image by detecting the edges of the object given by the selected channels. If no mean-center options are selected, then the object is not mean-centered.
%
\item[Resize Images] If the resize images option is selected, then each image is scaled/dialated so that the object occupies 80\% of the total image frame. This removes any size effects/variations in the images. 
%
\item[Number of Pixels] The number of pixels wanted in the images for analysis. This does {\em not} have to be the same number of pixels as the true resolution of the images. The images are assumed to be square. 
%
\end{description}

@@

\subsection*{Read Images}

\begin{par}
The first step is to read in images. This is done using the \texttt{read\_images} function For two-dimensional images, it requires that all the images are stored in a single directory, all with a consistent prefix, and indexed with two digits starting from \texttt{01}. For three-dimensional z-stack, it assumes each z-stack is stored in a single directory. All of the z-stack directories are required to begin with a consistent prefix, and be indexed with two digits starting from \texttt{01}. Similarly, all of the images are required to begin with a consistent prefix, and be indexed with two digits starting from \texttt{01} for each z-stack.
\end{par} \vspace{1em}
\begin{verbatim}
% directory where images are stored
image_dir = 'drosophila_fixed_images';

@@

% number of pixels
npixels = 100;

@@

% read in images
% images are stored in the variable images_raw
[images_raw, nchannels] = read_images(image_dir, image_name, image_ext, ...
    stack_name, nimages, nstack, npixels, dim);
\end{verbatim}

\subsection*{Show images}

@@

\includegraphics [width=4in]{sample_input_file_01.eps}

\subsection*{Preprocess Images}

\begin{par}
Now, we must preprocess the images before registration and ordering, to remove any imaging and/or experimental artifacts. We do this via the \texttt{apply\_image\_functions} function For this particular imaging data set, there are three channels. DAPI, a nuclear stain, is in the first (red) channel. dpERK is in the second (green) channel. Dl is in the third (blue) channel.
\end{par} \vspace{1em}
\begin{verbatim}
% channel weights
% because the nuclear signal is wider spread throughout the image, as well
% as noiser, we choose to scale the first channel by half
channel_weight = [0.5 1 1];

% we blur the image slightly (5%) so that small-scale effects of individual
% nuclei are smoothened
channel_blur = [0.05 0.05 0.05];

% because the absolute intensity of the nuclear signal is not informative
% or important (only the presence or absence of signal is important,
% indicating the presence or absence of nuclei), we normalize the first channel
% We do not normalize the other two channels, as changes in intensity are
% informative to the developmental processes.
channel_normalize = [1 0 0];

% we use the nuclear channel to mean center the images, as this channel
% parameterizes/captures the entire embryo
channel_mean_center = [1 0 0];

% we resize the images to all be (approximately) the same size, to remove
% any variations due to size effects, as size effects are known to not be
% important for this developmental system
resize_image = true;

% we then apply these image functions of normalization, blurring,
% reweighting, and mean-centering
images = apply_image_functions(images_raw, dim, channel_weight, ...
    channel_blur, channel_normalize, channel_mean_center, resize_image);

@@

\includegraphics [width=4in]{sample_input_file_02.eps}

\subsection*{Calculate pairwise alignments}

@@

\subsection*{Apply vector diffusion maps}

@@

\includegraphics [width=4in]{sample_input_file_03.eps}

>>>>

%\usepackage{epstopdf}

@@

All algorithms and analysis are implemented in MATLAB\textsuperscript{\textregistered} (R2013b, The MathWorks, Natick, Massachusetts).
%
The code requires MATLAB along with the Image Processing Toolbox.

@@

The code can be used either via an interactive GUI
or by writing a simple script.
%
The basic steps are as follows:
\begin{enumerate}
   \item Read in images
   \item Preprocess the images (smooth and/or equalize channels, mean-center, etc.)
   \item Run vector diffusion maps to register and order preprocessed images
   \item (Optional) Save registered and ordered images
   \item (Optional) Compute an average trajectory
\end{enumerate}

@@

Open ``register\_order\_gui.fig'', which is available in the downloaded code directory.
%
A window should appear (see below)

@@

@@

First populate the relevant parameters for where the images are stored.
%
The fields are as follows
\begin{description}
%
\item[2D/3D] Select ``2D'' if the images are traditional 2D images. Select ``3D'' if the images are z-stacks.
%
\item[Image Directory] Enter the directory where the images or z-stacks are stored. This directory can be an absolute path (e.g. \texttt{C:/Documents/Images}) or a path relative to the directory where this code is stored (e.g. \texttt{../Images})
%
\item[Image Prefix] The prefix for all image file names. The images are assumed to named as the image prefix, followed by a number. 
%
\item[Stack Prefix] (only for 3D z-stacks) The prefix for all stack folder names. Each stack is assumed to be stored in a single folder, with the folder's name as the stack prefix, followed by a number. 
%
\item[Image Extension] The extension of the images (e.g., tif, jpg)
%
\item[Number of Images] The number of images or z-stacks. Images are assumed to be stored as the image prefix, followed by a two-digit index, where the indices run from 1 to the number of images (e.g., image\_prefix01.tif, image\_prefix02.tif, etc.). For three-dimensional images, the z-stacks are assumed to be stored in individual folders whose names are the stack prefix, followed by a two-digit index, where the indices run from 1 to the number of stacks (e.g., z\_stack\_prefix01/, z\_stack\_prefix02/, etc.)
%
\item[Number of Images in Stack] (only for z-stacks) The number of images per z-stack. Images are assumed to be stored in the individual stack directories as the stack prefix, followed by a two-digit index, where the indices run from 1 to the number of images in the stack (e.g., stack\_prefix01.tif, stack\_prefix02.tif, etc.). 
%
\end{description}

@@

Define the relevant preprocessing operations. 
%
The options are as follows
%
\begin{description}
\item[Weight] Each channel in the images is scaled by the weight factor, where all weights at 1 does not adjust the channel intensities. For grayscale images, only the first weight is used. 
%
\item[Blur] Each channel in the images is blurred using a disc factor with radius equal to a fraction of the image as indicated by the ``blur'' field. For example, setting the blur to 5\% means that the image is blurred using a disc filter with radius $0.05 \times$ the number of pixels. 
%
\item[Normalize] If the normalize option is selected for a specific channel, then Contrast-Limited Adaptive Histogram Equalization is applied to that channel to normalize the intensities across the image. This is important for signals whose absolute intensity is not meaningful or informative. 
%
\item[Use to Mean-Center] If the mean-center option is selected for any of the channels, then the object is centered in the image by detecting the edges of the object given by the selected channels. If no mean-center options are selected, then the object is not mean-centered.
%
\item[Resize Images] If the resize images option is selected, then each image is scaled/dialated so that the object occupies 80\% of the total image frame. This removes any size effects/variations in the images. 
%
\item[Number of Pixels] The number of pixels wanted in the images for analysis. This does {\em not} have to be the same number of pixels as the true resolution of the images. The images are assumed to be square. 
%
\end{description}

@@

First, make sure that the directory which stores the downloaded code is on your MATLAB path, or that you are in that directory in MATLAB.

\begin{par}
The first step is to read in images, using the \texttt{read\_images} function. For two-dimensional images, it requires that all the images are stored in a single directory, all with a consistent prefix, and indexed with two digits starting from \texttt{01}. For three-dimensional z-stack, it assumes each z-stack is stored in a single directory. All of the z-stack directories are required to begin with a consistent prefix, and be indexed with two digits starting from \texttt{01}. Similarly, all of the images are required to begin with a consistent prefix, and be indexed with two digits starting from \texttt{01} for each z-stack.
\end{par} \vspace{1em}
\begin{verbatim}
% directory where images are stored
image_dir = 'drosophila_fixed_images';

@@

@@

% read in images
% images are stored in the variable images_raw
% nchannels stores the number of channels per image
[images_raw, nchannels] = read_images(image_dir, image_name, image_ext, ...
    stack_name, nimages, nstack, dim);
\end{verbatim}

\subsubsection{Show images}

@@

\includegraphics [width=4in]{sample_input_file_01.jpg}

\begin{par}
Now, we must preprocess the images using the \texttt{apply\_image\_functions} function before registration and ordering, to remove any imaging and/or experimental artifacts.
\end{par} \vspace{1em}
\begin{verbatim}
% number of pixels
% images will be reduced to npixels x npixels resolution
npixels = 100;

% channel weights
% we scale the first (red) channel by half, and keep the second (green) and
% third (blue) channels at their input values
channel_weight = [0.5 1 1];

% channel blur
% we blur each of the channels by 5%
channel_blur = [0.05 0.05 0.05];

% channel normalization
% we normalize the first (red) channel using histogram equalization
% we do not normalize the second (green) or third (blue) channels
channel_normalize = [1 0 0];

% mean-center
% we use the first (red) channel to detect the edges of the object in order
% to mean center the object
channel_mean_center = [1 0 0];

% resize
% we choose to resize the images so all objects are (approximately)
% the same size, to remove any variations due to size effects
resize_image = true;

% we then apply these image functions of normalization, blurring,
% reweighting, and mean-centering
images = apply_image_functions(images_raw, npixels, dim, channel_weight, ...
    channel_blur, channel_normalize, channel_mean_center, resize_image);

@@

\includegraphics [width=4in]{sample_input_file_02.jpg}

\subsubsection{Calculate pairwise alignments}

@@

\subsubsection{Apply vector diffusion maps}

@@

\includegraphics [width=4in]{sample_input_file_03.jpg}

\subsubsection{Calculate average trajectory}

\begin{par}
We can calculate an average trajectory from our set of (registered and ordered) images
\end{par} \vspace{1em}
\begin{verbatim}
% nsubimages is the desired number of images in the average trajectory
nsubimages = 10;

% avg_width is the (approximate) width of the averaging window used to
% compute each of the images in the average trajectory
avg_width = 4;

% compute the average trajectory
avg_images = compute_average_trajectory(images_analyzed, nsubimages, avg_width);

% plot the images (optional)
plot_images(avg_images, dim)
\end{verbatim}

\includegraphics [width=4in]{sample_input_file_04.jpg}

\section{Parameters for Data Sets}

The four data sets used in the paper are available with the code in the relevant subdirectories. The parameters used (as defined in the script in the previous section) are defined below.

\subsection{{\em Drosophila} live imaging}

\begin{verbatim}

image_dir = 'drosophila_live_imaging/expt01';
image_name = 'image';
image_ext = 'tif';
stack_name = '';
nimages = 40;
nstack = 0;
dim = 2;

npixels = 100;
channel_weight = 1;
channel_blur = 0.05;
channel_normalize = 1;
channel_mean_center = 0;
resize_image = false;

ang_dis = 10;

eps_scale = 0.25;

\end{verbatim}

\subsection{Zebrafish}

\begin{verbatim}

image_dir = 'zebrafish';
image_name = 'image';
image_ext = 'tif';
stack_name = '';
nimages = 120;
nstack = 0;
dim = 2;

npixels = 100;
channel_weight = 1;
channel_blur = 0;
channel_normalize = 0;
channel_mean_center = 1;
resize_image = false;

ang_dis = 10;

eps_scale = 0.25;

\end{verbatim}

\subsection{{\em Drosophila} fixed images}

\begin{verbatim}

image_dir = 'drosophila_fixed_images';
image_name = 'emb';
image_ext = 'tif';
stack_name = '';
nimages = 120;
nstack = 0;
dim = 2;

npixels = 100;
channel_weight = [0.5 1 1];
channel_blur = [0.05 0.05 0.05];
channel_normalize = [1 0 0];
channel_mean_center = [1 0 0];
resize_image = true;

ang_dis = 10;

eps_scale = 0.25;

\end{verbatim}

\subsection{Wing disc}

\begin{verbatim}

image_dir = 'wing_disc_stacks';
image_name = 'disc';
image_ext = 'tif';
stack_name = 'z';
nimages = 46;
nstack = 21;
dim = 3;

npixels = 100;
channel_weight = [1 1 1];
channel_blur = [0 0 0];
channel_normalize = [0 0 0];
channel_mean_center = [0 0 1];
resize_image = false;

ang_dis = 10;

eps_scale = 0.5;
\end{verbatim}

\section{Parameter Values and Guidance}

\begin{description}

\item[Weight] The relative weights of the channels in multichannel images in important for computing meaningful distances for diffusion maps. The total intensity for each channel should be of the same order if all channels are to be considered as ``equally important.'' If the channels have dramatically different intensities in the raw images, they should be adjusted via the weight functions.
%
\item[Blur] The image should be blurred to remove fine-scale noise and variations. Typically, blur of 5\%-10\% are appropriate for removing such variability. In some cases, blurring is not necessary to recover an accurate registration and ordering. 
%
\item[Normalize] Channels should be normalized if the absolute intensity of a specific signal is not meaningful. For example, DAPI, a DNA marker, identifies the nuclei in the {\em Drosophila} embryo, but the absolute intensity of the signal is not important. 
%
\item[Use to Mean-Center] Objects should be (approximately) mean-centered in the frame prior to registration. The channel(s) that best identify/delineate the periphery of the object should be selected for use in the mean-centering procedure.
%
\item[Resize Images] Object should be resized if size differences are not meaningful/informative about the underlying developmental dynamics. 
%
\item[Number of Pixels] The image resolution should be chosen as low as possible such that the relevant developmental structures and patterns are still discernible/evident. A higher image resolution will result in a higher computational cost. 
%
\item[Kernel Scale] The kernel scale defines how similar two images have to be to be considered developmentally close. Small values of the kernel will preserve finer scale patternds and structure in the ordering algorihtm, but will also be more sensitive to noise. We have found values of 0.25-0.5 to often be suitable for our studies. Noisy data with a lower temporal resolution (such as the wing disc data) often requires a larger value of the kernel scale than clean data with a high temporal resolution (such as the zebrafish live imaging data). 
%
\item[Angular Discretization] The angular discretization sets the resolution for computing the pairwise alignments. For example, if the angular discretization is 10$^\circ$, then the pairwise alignments are computed by computing the image discrepancy after 10$^\circ$ rotation, after a 20$^\circ$ rotation, ..., up to a 360$^\circ$ rotation, and then selecting the rotation which minimizes the discrepancy. Lower values of the angular discretization increase the accuracy of the method, but also increase the computational cost. We have found that $10^\circ$ is often a good value of the angular discretization, which balances computational cost and accuracy.

\end{description}

### 2015-02-02 image_analysis_paper_development2/code_for_distribution/README/sample_input_file.tex
updated report

\begin{itemize}
\setlength{\itemsep}{-1ex}
   \item Basic Algorithm
   \item Read Images
   \item Show images
   \item Preprocess Images
   \item Calculate pairwise alignments
   \item Apply vector diffusion maps
\end{itemize}
\begin{verbatim}
clear all
close all
\end{verbatim}

@@

% read in images
% images are stored in the variable images_raw
[images_raw, nchannels] = read_images(image_dir, image_name, image_ext, ...
    stack_name, nimages, nstack, npixels, dim);
\end{verbatim}

@@

\includegraphics [width=4in]{sample_input_file_01.eps}

@@

\begin{par}
Now, we must preprocess the images before registration and ordering, to remove any imaging and/or experimental artifacts. We do this via the \texttt{apply\_image\_functions} function For this particular imaging data set, there are three channels. DAPI, a nuclear stain, is in the first (red) channel. dpERK is in the second (green) channel. Dl is in the third (blue) channel.
\end{par} \vspace{1em}
\begin{verbatim}
% channel weights
% because the nuclear signal is wider spread throughout the image, as well
% as noiser, we choose to scale the first channel by half
channel_weight = [0.5 1 1];

% we blur the image slightly (5%) so that small-scale effects of individual
% nuclei are smoothened
channel_blur = [0.05 0.05 0.05];

% because the absolute intensity of the nuclear signal is not informative
% or important (only the presence or absence of signal is important,
% indicating the presence or absence of nuclei), we normalize the first channel
% We do not normalize the other two channels, as changes in intensity are
% informative to the developmental processes.
channel_normalize = [1 0 0];

% we use the nuclear channel to mean center the images, as this channel
% parameterizes/captures the entire embryo
channel_mean_center = [1 0 0];

% we resize the images to all be (approximately) the same size, to remove
% any variations due to size effects, as size effects are known to not be
% important for this developmental system
resize_image = true;

% we then apply these image functions of normalization, blurring,
% reweighting, and mean-centering
images = apply_image_functions(images_raw, dim, channel_weight, ...
    channel_blur, channel_normalize, channel_mean_center, resize_image);

@@

\includegraphics [width=4in]{sample_input_file_02.eps}

@@

\includegraphics [width=4in]{sample_input_file_03.eps}

>>>>

\begin{itemize}
\setlength{\itemsep}{-1ex}
   \item Basic Algorithm
   \item Read Images
   \item Show images
   \item Preprocess Images
   \item Calculate pairwise alignments
   \item Apply vector diffusion maps
   \item Calculate average trajectory
\end{itemize}
\begin{verbatim}
clear all
close all
\end{verbatim}

@@

% read in images
% images are stored in the variable images_raw
% nchannels stores the number of channels per image
[images_raw, nchannels] = read_images(image_dir, image_name, image_ext, ...
    stack_name, nimages, nstack, dim);
\end{verbatim}

@@

\includegraphics [width=4in]{sample_input_file_01.jpg}

@@

\begin{par}
Now, we must preprocess the images using the \texttt{apply\_image\_functions} function before registration and ordering, to remove any imaging and/or experimental artifacts.
\end{par} \vspace{1em}
\begin{verbatim}
% number of pixels
% images will be reduced to npixels x npixels resolution
npixels = 100;

% channel weights
% we scale the first (red) channel by half, and keep the second (green) and
% third (blue) channels at their input values
channel_weight = [0.5 1 1];

% channel blur
% we blur each of the channels by 5%
channel_blur = [0.05 0.05 0.05];

% channel normalization
% we normalize the first (red) channel using histogram equalization
% we do not normalize the second (green) or third (blue) channels
channel_normalize = [1 0 0];

% mean-center
% we use the first (red) channel to detect the edges of the object in order
% to mean center the object
channel_mean_center = [1 0 0];

% resize
% we choose to resize the images so all objects are (approximately)
% the same size, to remove any variations due to size effects
resize_image = true;

% we then apply these image functions of normalization, blurring,
% reweighting, and mean-centering
images = apply_image_functions(images_raw, npixels, dim, channel_weight, ...
    channel_blur, channel_normalize, channel_mean_center, resize_image);

@@

\includegraphics [width=4in]{sample_input_file_02.jpg}

@@

\includegraphics [width=4in]{sample_input_file_03.jpg}

\subsection*{Calculate average trajectory}

\begin{par}
We can calculate an average trajectory from our set of (registered and ordered) images
\end{par} \vspace{1em}
\begin{verbatim}
% nsubimages is the desired number of images in the average trajectory
nsubimages = 10;

% avg_width is the (approximate) width of the averaging window used to
% compute each of the images in the average trajectory
avg_width = 4;

% compute the average trajectory
avg_images = compute_average_trajectory(images_analyzed, nsubimages, avg_width);

% plot the images (optional)
plot_images(avg_images, dim)
\end{verbatim}

\includegraphics [width=4in]{sample_input_file_04.jpg}
